Turbo fast no compromise gravitational wave parameter estimation

### 2022-12-06 src/tex/ms.tex
Shuffle position of heterodyne likelihood
update draft
Add some discussion around the comparison between our work and others

In this section, we provide a number of sanity checks as well as a few real life applications

@@

@@

\kw{GW159014}
\kw{GW170817}

@@

@@

% Compared to other works

There are recent works from different groups on speeding up parameter estimation
of gravitational wave, including efficient reparameterization and deep learning
\kw{cite}. While all of these methods are pushing down minutes-scale parameter
estimation with high fidelity, we would like to highlight the unique strength of
this work. Compared to \kw{Stephan's paper}, we do not rely on pre-training the
neural network on a large collection of waveforms and noise realization. This
means as soon as new waveforms models and noise models are available, our
algorithm can already be deployed. Furthermore, our method is essentially an
MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
are only using the normalizing flow as a proposal distribution, and the
normalizing flow is trained jointly with a local sampler, we do not suffer from
overfitting since our data is always approaching the target distribution, in
this sense, we do not introduce potential extra systematic error to inference
result. Compared to \kw{Tousif's paper}, we do not rely on handcrafted
reparameterization of the coordinate systems. If the reparameterization scheme
is known ahead of time, a nice reparameterization is also encouraged. However,
handcrafted reparameterization depends on the assumptions used in deriving the
reparameterization scheme, which would inevitably run into limitations of use
cases. Our work can be viewed as an automatic reparameterization powered by
normalizing flow, while not being exact hence as efficient as an explicit
reparameterization, the method proposed in this work is applicable to a much
boarder class of problems, such as parameter estimation with precessing
waveforms, testing GR, and multi-event joint inference.

>>>>

In this section, we provide a number of sanity checks as well as a number of real life usecases

@@

In order to verify the fidelity of the sampler, we can inject a signal
into the parameter estimation pipeline and compared the recovery parameters to
the true parameters.

% Show injection pp plot

@@

To demonstrate the performance of our parameter estimation pipeline, we apply
our pipeline to analysis GW150914 and GW170817.

\kw{GW159014}

\kw{GW170817}

@@

In this work, we present a PE pipeline for GW events that is efficient and flexible.
Our implemneted the following major features to achieve its performance:
\begin{enumerate}
\item Differential waveform models
\item Normalizing flow enhanced MCMC sampler
\item Heterodyned likelihood
\item Native support for accelerator
\item JIT compilation
\end{enumerate}

% Compared to other works
There are recent works from different groups on speeding up parameter estimation
of gravitational wave, including efficient reparameterization and deep learning
\kw{cite}. While all of these methods are able to get down to minutes-scale parameter
estimation with high fidelity, we would like to highlight the unique strength of
this work, and potential interplay between our work and other's work.
% Stephen's work
Compared to \kw{Stephan's paper}, we do not rely on pre-training the
neural network on a large collection of waveforms and noise realization. This
means as soon as new waveforms models and noise models are available, our
algorithm can already be deployed. Furthermore, our method is essentially an
MCMC algorithm, which inherits the merit of convergence measures in MCMC. As we
are only using the normalizing flow as a proposal distribution, and the
normalizing flow is trained jointly with a local sampler, we do not suffer from
overfitting since our data is always approaching the target distribution, in
this sense, we do not introduce potential extra systematic error to inference
result.

While our pipeline use the sample generated by the local sampler for training,
one can also supply a pre-trained normalizing flow model to our pipeline to
bypass the training stage. This can further reduce the total runtime of our
inference pipeline. However, this may introduce potential systematic bias in the
inference result if the pre-trained network is not able to capture the
complexity presented in the data.

%Tousif's work
Compared to \kw{Tousif's paper}, we do not rely on handcrafted
reparameterization of the coordinate systems. If the reparameterization scheme
is known ahead of time, a nice reparameterization is also encouraged. However,
handcrafted reparameterization depends on the assumptions used in deriving the
reparameterization scheme, which would inevitably run into limitations of use
cases. Our work can be viewed as an automatic reparameterization powered by
normalizing flow, while not being exact hence as efficient as an explicit
reparameterization, the method proposed in this work is applicable to a much
boarder class of problems, such as parameter estimation with precessing
waveforms, testing GR, and multi-event joint inference.

It is always beneficial to reparameterize if the reparameterization is known
ahead-of-time. For the class of problems where \kw{Tousif's paper} is applicable
to, we can incorporate the reparameterization scheme into our MCMC pipeline to
reduce the complexity of the problem, hence speeding up the training phase.
On the other hand, if there are part of the target posterior that are not
included in the reparameterization scheme, the normalizing flow should still be
able to learn to produce accurate samples efficiently.

As there are multiple group

@@
